---
title: "Summary of `tidyjson`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is it?

A package by [Jeremy Stanley](https://github.com/jeremystan) that provides "(a)n easy and consistent way to turn JSON into tidy data frames that are natural to work with in `dplyr`, `ggplot2` and other tools."

## Main functions

### Data Exploration

- determine_types 
- enter_object
- read_json 
- json_lengths 
- json_types
- [.tbl_json 

### Data Manipulation

- gather_array 
- gather_keys 
- my_unlist
- replace_nulls 
- spread_values
- as.tbl_json 

### Data Creation

- tbl_json
- append_values 
- append_values_factory 
- append_values_type

### Other

- allowed_json_types 
- jfactory  
- jfunctions  
- list_path  
- prep_path  
- wrap_dplyr_verb  

## Classes 

- `tbl_json` 

## Data Sets

- `commits` - "Commit data for the dplyr repo from github API"
- `companies ` - [1,000 Startup Companies' info](http://jsonstudio.com/resources/)
- `issues` - "Issue data for the dplyr repo from github API"
- `worldbank` - [Projects funded by the World Bank](http://jsonstudio.com/resources/)

## Walk-through Example 1: US Zip Codes

1. Read in data - `tidyjson` doesn't have a way to read newline delimited JSON data into R. Use `ndjson` instead.

```{r readzip}
zips <- ndjson::stream_in("data/zips.json")
dplyr::glimpse(zips)
```

Okay, that was easy. Maybe too easy... Let's move on to another example.

## Walk-through Example 2: World Bank

1. Get data - The data are already a part of the package!

```{r wbread}
jsonbikes <- httr::GET("http://citibikenyc.com/stations/json")
bikes <- rawToChar(jsonbikes$content)
saveRDS(bikes, "bikes.json")
read_json("bikes.json")
```

2. Explore the World Bank 

- determine_types
```{r wbconvert}
```
- enter_object
- read_json 
- json_lengths 
- json_types
- [.tbl_json 


3. Start restructuring the character JSON data to a nice data frame structure.

```{r echo=FALSE, eval=FALSE}
#dump from another file
# learning about tidyjson
library(dplyr)
purch_json <- '
[
  {
  "name": "bob", 
  "purchases": [
  {
  "date": "2014/09/13",
  "items": [
  {"name": "shoes", "price": 187},
  {"name": "belt", "price": 35}
  ]
  }
  ]
  },
  {
  "name": "susan", 
  "purchases": [
  {
  "date": "2014/10/01",
  "items": [
  {"name": "dress", "price": 58},
  {"name": "bag", "price": 118}
  ]
  },
  {
  "date": "2015/01/03",
  "items": [
  {"name": "shoes", "price": 115}
  ]
  }
  ]
  }
  ]'
  
  library(tidyjson)
  
  purch_items <- purch_json %>%
    gather_array %>%                                     # stack the users 
    spread_values(person = jstring("name")) %>%          # extract the user name
    enter_object("purchases") %>% gather_array %>%       # stack the purchases
    spread_values(purchase.date = jstring("date")) %>%   # extract the purchase date
    enter_object("items") %>% gather_array %>%           # stack the items
    spread_values(                                       # extract item name and price
      item.name = jstring("name"),
      item.price = jnumber("price")
    ) %>%
    select(person, purchase.date, item.name, item.price)
  
  (this.raw.content %>% 
      gather_keys() %>% json_types())[2,] %>% 
    gather_array() %>% gather_keys %>% 
    spread_values(npi = jstring("npi")) %>% 
    enter_object("practices")  %>% gather_keys
  
  (this.raw.content %>% 
      gather_keys() %>% json_types())[2,] %>% 
    gather_array() %>% 
    spread_values(npi = jstring("npi")) %>% 
    enter_object("practices") %>% gather_array %>%
    spread_values(loc_slug = jstring("location_slug"),
                  dist = jstring("distance"), 
                  lat = jstring("lat"), lon = jstring("lon"), 
                  phones = jstring("phones")
    )
```